---
title: 'Analysis of the Social and Political Factors for Success/Failure in Tackling
  the Covid-19 Pandemic  '
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

                      ####LIBRARIES####
library(tidyverse)
library(tidyr)
library(ggplot2)
####DATASETS####
dataset1 <- read.csv("https://raw.githubusercontent.com/QMUL-SPIR/Public_files/master/datasets/Covid2020.csv")
dataset_omit <- read.csv("https://raw.githubusercontent.com/QMUL-SPIR/Public_files/master/datasets/Covid2020.csv")
dataset_omit <- drop_na(dataset_omit, extreme_poverty)
```
## Introduction

what factors lead to the best or worst responses to the Covid –19 crisis? Is it wealth, who leads the country? Maybe how the country is run? Maybe it's something out of their control. I have been given four hypotheses to test and using data on an assortment of countries from around the world, I will test them and see how these factors affect their death toll. The theories are as follows:

* Richer Countries do better than Poorer countries. 

* Countries led by women are more successful than those led by men. 

* The success in dealing with the pandemic is related to the proportion of older people living in the country. 

* Democratic countries are more successful than authoritarian countries. 

To test these theories, I collected and analyzed data from various countries around the world and used a series of measurements and characteristics to determine what effects the a country's success at tackling the COVID-19 pandemic. 

## Justification:
First, I must justify my initial hypothesis, what I believe the data will show me using prior research from sources, like the news, academic papers etc. (to prove that I didn't pull these statements out of thin air!). 

1. Richer Countries do better than Poorer countries.

A country with a higher GDP per capita and lower extreme poverty levels would be more likely to have a robust healthcare system and greater investment into pandemic preparation.



2. Countries led by women are more successful than those led by men. 

In the early days of the pandemic, the media highlighted the success of New Zealand Prime Minister Jacinda Ardern and other female led nations around the world at having lower death rates compared to their male counterparts.

* https://www.theguardian.com/world/2020/aug/18/female-led-countries-handled-coronavirus-better-study-jacinda-ardern-angela-merkel

* https://www.ft.com/content/6b597385-ba51-413a-96bd-cb75d3446718

* https://www.forbes.com/sites/avivahwittenbergcox/2020/04/13/what-do-countries-with-the-best-coronavirus-reponses-have-in-common-women-leaders/

3. The success in dealing with the pandemic is related to the proportion of older people living in the country. 

The death rate of COVID-19 infected individuals is much higher in those aged above 60. In nations with a higher older population such as Japan, the government would be prepared to take stronger measures to protect the most vulnerable as they could expect more deaths in their population.

4.Democratic countries are more successful than authoritarian countries. 

## The Results:

Variable                                            Name of the Relevant MCT    Value of the MCT                    Name of the Relevant MD                   Value of the MD 
---------                                           -------------------------   -----------------                   ------------------------                  ----------------
Total Deaths Per Million                            Continuous                  140.40                              variance and/or standard deviation        209.88    
GDP Per Capita                                      Continuous                  14910.69                            variance and/or standard deviation        16071.77
Female Head of Government                           Nominal                     0 (Male) = 107  1 (Female) = 9      proportion in each category               0 (Male) = 0.92  1 (Female) = 0.08
V-Dem Polyarchy                                     Continuous                  0.56                                variance and/or standard deviation        0.22



#### Richer Countries do better than Poorer countries.
```{r, echo=FALSE}
gg_gdp_per_cap <- ggplot(data = dataset_omit, 
                         mapping = aes(x = gdp_per_capita  , y = total_deaths_per_million)) +
  geom_text(size = 2, 
            colour = "red", 
            aes(label = location))

gg_gdp_per_cap + geom_point()

```


```{r, echo=FALSE}
#gdp per capita

#GDP per capita binary
median_gdp_p_c <- median(dataset_omit$gdp_per_capita)

gdp_oneOrZero <- dataset_omit %>%
  mutate(
    gdp_oneOrZero_Median =
      factor(case_when(
        gdp_per_capita > median_gdp_p_c ~ 1,
        gdp_per_capita<= median_gdp_p_c ~ 0)))


gdp_1or0_dist <- ggplot(data = gdp_oneOrZero, aes(total_deaths_per_million, group = gdp_oneOrZero_Median)) +
geom_density(aes(color = gdp_oneOrZero_Median)) +
labs(x = "Total Deaths Per Million",
     y = "Density",
title = "Distribution of Deaths per Million Conditional on GDP per Capita") + # title
scale_color_discrete(name = "GDP per Capita",
labels = c("Below or equal median", "Above median")) + theme_minimal()

gdp_1or0_dist
```

```{r, echo=FALSE}
#GDP per capita binary
median_gdp_p_c <- median(dataset_omit$gdp_per_capita)

gdp_oneOrZero <- dataset_omit %>%
  mutate(
    gdp_oneOrZero_Median =
      factor(case_when(
        gdp_per_capita > median_gdp_p_c ~ 1,
        gdp_per_capita<= median_gdp_p_c ~ 0)))

t.test(total_deaths_per_million ~ gdp_oneOrZero_Median ,
       data = gdp_oneOrZero ,
       mu = 0,
       alt = "two.sided",
       conf = 0.95)
```

#### Countries led by women are more successful than those led by men.  

```{r, echo=FALSE}

#Female HoG Deaths
gg_female_HoG_deaths <- ggplot(data = dataset_omit, aes(x = total_deaths_per_million)) +
  geom_histogram(bins = 30)

geom_text(size = 2, 
          colour = "red", 
          aes(label = location))

gg_female_HoG_deaths + facet_wrap(~ female_HoG)
```



#### The success in dealing with the pandemic is related to the proportion of older people living in the country.

```{r, echo=FALSE}
#Population over 65

gg_pop_65_plus <- ggplot(data = dataset_omit, 
                         mapping = aes(x = aged_65_older , y = total_deaths_per_million)) +   
  geom_text(size = 2, 
  colour = "red", 
  aes(label = location))

gg_pop_65_plus + geom_point()
```


```{r, echo=FALSE}



#aged 65 plus binary
median_sixtyFiveBinary <- median(dataset_omit$aged_65_older)

sixtyFive_Median <- dataset_omit %>%
  mutate(
    SixtyFive_Median_Median =
      factor(case_when(
        aged_65_older > median_sixtyFiveBinary ~ 1,
        aged_65_older<= median_sixtyFiveBinary ~ 0)))

#aged 65 plus binary

aged65_dist <- ggplot(data = sixtyFive_Median, aes(total_deaths_per_million, group = SixtyFive_Median_Median)) +
  geom_density(aes(color = SixtyFive_Median_Median)) +
  labs(x = "Aged 65 or Over",
       y = "Density",
       title = "Distribution of Deaths per Million Conditional on Population aged 65 or Older") + # title
  scale_color_discrete(name = "GDP per Capita",
                       labels = c("Below or equal median", "Above median")) + theme_minimal()

aged65_dist
```

```{r, echo=FALSE}
t.test(total_deaths_per_million ~ SixtyFive_Median_Median ,
       data = sixtyFive_Median,
       mu = 0,
       alt = "two.sided",
       conf = 0.95)
```

#### Democratic countries are more successful than authoritarian countries.

```{r, echo=FALSE}
gg_vdem <- ggplot(data = dataset_omit, 
                         mapping = aes(x = vdem_polyarchy , y = total_deaths_per_million)) +   
  geom_text(size = 2, 
            colour = "red", 
            aes(label = location))

gg_vdem + geom_point()
```
```{r, echo=FALSE}
#vdem polyarchy binary

#vdem polyarchy binary 
median_vdemBinary <- median(dataset_omit$vdem_polyarchy)

Vdem_Median <- dataset_omit %>%
  mutate(
    Vdem_Median_Median =
      factor(case_when(
        vdem_polyarchy > median_vdemBinary ~ 1,
        vdem_polyarchy<= median_vdemBinary ~ 0)))



vdem_dist <- ggplot(data = Vdem_Median, aes(total_deaths_per_million, group = Vdem_Median_Median)) +
  geom_density(aes(color = Vdem_Median_Median)) +
  labs(x = "Vdem Polyarchy",
       y = "Density",
       title = "Distribution of Deaths per Million Conditional on Vdem Polyarchy") + # title
  scale_color_discrete(name = "GDP per Capita",
                       labels = c("Below or equal median", "Above median")) + theme_minimal()

vdem_dist
```


```{r, echo=FALSE}
# vdem polyarchy 

t.test(total_deaths_per_million ~ Vdem_Median_Median  ,
       data = Vdem_Median,
       mu = 0,
       alt = "two.sided",
       conf = 0.95)
```

## Conclusion: 

## Code:

```{r eval=TRUE}
# This is where the finished code will go. 

```